---
output: 
    word_document:
        reference_docx: document_template.docx
---

## 2016 District EOC Results

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggthemes)

system_base <- readstata13::read.dta13("K:/ORP_accountability/projects/2016_state_results/system_base_with_super_subgroup_2016.dta")

system_numeric <- readstata13::read.dta13("K:/ORP_accountability/projects/2016_state_results/system_numeric_with_super_subgroup_2016.dta") %>%
    rename(District = system_name) %>%
    mutate(District = sub("Special School District", "SSD", District),
        Math = ifelse(system %in% c(30, 60, 80, 100, 110, 140, 150, 190, 440, 580, 590, 710, 800, 821, 850, 890, 930), "Integrated Math", "Algebra")) 

state_numeric <- readstata13::read.dta13("K:/ORP_accountability/projects/2016_state_results/state_numeric_with_super_subgroup_2016.dta") %>%
    rename(District = system_name) %>%
    mutate(District = sub("Special School District", "SSD", District))
```

**Participation Rate**

For accountability purposes, districts are required to meet a 95 percent participation rate for all subject/subgroup combinations with 30 or more enrolled students.

The state participation rate was `r state_numeric[state_numeric$subject == "HS English" & state_numeric$subgroup == "All Students", ]$participation_rate` percent for High School English, and `r state_numeric[state_numeric$subject == "HS Math" & state_numeric$subgroup == "All Students", ]$participation_rate` percent for High School Math, and `r state_numeric[state_numeric$subject == "HS Science" & state_numeric$subgroup == "All Students", ]$participation_rate` percent for High School Science.

`r with(system_numeric, sum(subject == "HS English" & subgroup == "All Students" & (enrolled + enrolled_part_1_only + enrolled_part_2_only + enrolled_both) > 30 & participation_rate >= 95))` out of `r with(system_numeric, sum(subject == "HS English" & subgroup == "All Students" & (enrolled + enrolled_part_1_only + enrolled_part_2_only + enrolled_both) > 30))` eligible districts met a 95 percent participation rate in High School English.

``` {r, fig.width = 6}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS English") %>%
    filter(enrolled + enrolled_part_1_only + enrolled_part_2_only + enrolled_both >= 30) %>%
    ggplot(aes(x = participation_rate)) +
    geom_histogram(breaks = seq(75, 100, by = 2.5), col = "black", fill = "blue", alpha = 0.3) +
    theme_hc() +
    labs(x = "HS English Participation Rate", y = "District Count")
```

Districts which failed to meet a 95 percent participation rate for English:

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS English") %>%
    filter(enrolled + enrolled_part_1_only + enrolled_part_2_only + enrolled_both >= 30) %>%
    filter(participation_rate < 95) %>%
    arrange(participation_rate) %>%
    select(District, participation_rate) %>%
    rename(`Participation Rate` = participation_rate) %>%
    knitr::kable()
```

&nbsp;

`r with(system_numeric, sum(subject == "HS Math" & subgroup == "All Students" & (enrolled + enrolled_part_1_only + enrolled_part_2_only + enrolled_both) > 30 & participation_rate >= 95))` out of `r with(system_numeric, sum(subject == "HS Math" & subgroup == "All Students" & (enrolled + enrolled_part_1_only + enrolled_part_2_only + enrolled_both) > 30))` eligible districts met a 95 percent participation rate in High School Math.

``` {r, fig.width = 6}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS Math") %>%
    filter(enrolled + enrolled_part_1_only + enrolled_part_2_only + enrolled_both >= 30) %>%
    ggplot(aes(x = participation_rate)) +
    geom_histogram(breaks = seq(75, 100, by = 2.5), col = "black", fill = "blue", alpha = 0.3) +
    theme_hc() +
    labs(x = "HS Math Participation Rate", y = "District Count")
```

Districts which failed to meet a 95 percent participation rate for math:

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS Math") %>%
    filter(enrolled + enrolled_part_1_only + enrolled_part_2_only + enrolled_both >= 30) %>%
    filter(participation_rate < 95) %>%
    arrange(participation_rate) %>%
    select(District, participation_rate) %>%
    rename(`Participation Rate` = participation_rate) %>%
    knitr::kable()
```

&nbsp;

**All Students Achievement**

The percentage of students On Track/Mastered in Math ranged from `r system_numeric %>% filter(subgroup == "All Students" & subject == "HS Math" & valid_tests >= 30) %>% with(min(pct_on_mastered))` to `r system_numeric %>% filter(subgroup == "All Students" & subject == "HS Math" & valid_tests >= 30) %>% with(max(pct_on_mastered))`.

``` {r, fig.width = 6}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS Math") %>%
    filter(valid_tests >= 30) %>%
    ggplot(aes(x = pct_on_mastered)) + 
    geom_histogram(breaks = seq(0, 60, by = 5), col = "black", fill = "blue", alpha = 0.3) + 
    theme_hc() + 
    labs(x = "HS Math Percent On Track/Mastered", y = "District Count")
```

&nbsp;

The highest performing districts in Math were...

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS Math") %>%
    filter(valid_tests >= 30) %>%
    arrange(desc(pct_on_mastered)) %>%
    select(District, pct_on_mastered) %>%
    rename(`Percent On Track/Mastered` = pct_on_mastered) %>%
    slice(1:10) %>%
    knitr::kable()
```

&nbsp;

The percentage of students On Track/Mastered in English ranged from `r system_numeric %>% filter(subgroup == "All Students" & subject == "HS English" & valid_tests >= 30) %>% with(min(pct_on_mastered))` to `r system_numeric %>% filter(subgroup == "All Students" & subject == "HS English" & valid_tests >= 30) %>% with(max(pct_on_mastered))`.

``` {r, fig.width = 6}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS English") %>%
    filter(valid_tests >= 30) %>%
    ggplot(aes(x = pct_on_mastered)) + 
    geom_histogram(breaks = seq(0, 60, by = 5), col = "black", fill = "blue", alpha = 0.3) + 
    theme_hc() + 
    labs(x = "HS English Percent On Track/Mastered", y = "District Count")
```

&nbsp;

The highest performing districts in English were...

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS English") %>%
    filter(valid_tests >= 30) %>%
    arrange(desc(pct_on_mastered)) %>%
    select(District, pct_on_mastered) %>%
    rename(`Percent On Track/Mastered` = pct_on_mastered) %>%
    slice(1:10) %>%
    knitr::kable()
```

&nbsp;

Most districts performed better in English than in Math. Higher performing districts in English also tended to be higher performing districts in Math.

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject %in% c("HS Math", "HS English")) %>%
    filter(valid_tests >= 30) %>%
    select(District, subject, pct_on_mastered, Math) %>%
    spread(subject, pct_on_mastered) %>%
    ggplot(aes(x = `HS Math`, y = `HS English`, colour = Math)) +
    geom_point() + 
    geom_abline(intercept = 0, slope = 1, color = "red") +
    theme_hc() +
    labs(x = "HS Math Percent On Track/Mastered", y = "HS English Percent On Track/Mastered")
```

**All Students Relative Performance**

Changes in Math relative performance... (Some large swings)

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS Math") %>%
    ggplot(aes(x = PA_percentile_prior, y = OM_percentile, colour = Math)) +
    geom_point() +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    theme_hc() +
    labs(x = "2015 Math Proficient/Advanced Percentile", y = "2016 Math On Track/Mastered Percentile")
```

&nbsp;

Districts with the largest increases in relative performance in Math were...

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS Math") %>%
    mutate(`Change in Percentile Rank` = OM_percentile - PA_percentile_prior) %>%
    arrange(desc(`Change in Percentile Rank`)) %>%
    select(District, OM_percentile, PA_percentile_prior, `Change in Percentile Rank`) %>%
    rename(`2016 Percentile Rank` = OM_percentile, `2015 Percentile Rank` = PA_percentile_prior) %>%
    slice(1:10) %>%
    knitr::kable()
```

&nbsp;

Districts with the largest decreases in relative performance in Math were...

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS Math") %>%
    mutate(`Change in Percentile Rank` = OM_percentile - PA_percentile_prior) %>%
    arrange(`Change in Percentile Rank`) %>%
    select(District, OM_percentile, PA_percentile_prior, `Change in Percentile Rank`) %>%
    rename(`2016 Percentile Rank` = OM_percentile, `2015 Percentile Rank` = PA_percentile_prior) %>%
    slice(1:10) %>%
    knitr::kable()
```

&nbsp;

Changes in English relative performance... (Not as big swings)

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS English") %>%
    ggplot(aes(x = PA_percentile_prior, y = OM_percentile)) +
    geom_point() +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    theme_hc() +
    labs(x = "2015 English Proficient/Advanced Percentile", y = "2016 English On Track/Mastered Percentile")
```

&nbsp;

Districts with the largest increases in relative performance in English were...

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS English") %>%
    mutate(`Change in Percentile Rank` = OM_percentile - PA_percentile_prior) %>%
    arrange(desc(`Change in Percentile Rank`)) %>%
    select(District, OM_percentile, PA_percentile_prior, `Change in Percentile Rank`) %>%
    rename(`2016 Percentile Rank` = OM_percentile, `2015 Percentile Rank` = PA_percentile_prior) %>%
    slice(1:10) %>%
    knitr::kable()
```

&nbsp;

Districts with the largest decreases in relative performance in English were...

``` {r}
system_numeric %>%
    filter(subgroup == "All Students" & subject == "HS English") %>%
    mutate(`Change in Percentile Rank` = OM_percentile - PA_percentile_prior) %>%
    arrange(`Change in Percentile Rank`) %>%
    select(District, OM_percentile, PA_percentile_prior, `Change in Percentile Rank`) %>%
    rename(`2016 Percentile Rank` = OM_percentile, `2015 Percentile Rank` = PA_percentile_prior) %>%
    slice(1:10) %>%
    knitr::kable()
```

&nbsp;

**Subgroup Performance**

Performance in HS Math and English by subgroup... (Consistent with above, English looks better than math)

``` {r, fig.width = 7, fig.height = 7}
system_numeric %>%
    filter(!(subgroup %in% c("All Students", "Super Subgroup"))) %>%
    filter(valid_tests >= 30) %>%
    ggplot(aes(x = pct_on_mastered)) +
    geom_histogram(col = "black", fill = "blue", alpha = 0.3) +
    facet_grid(subject ~ subgroup) + 
    theme_hc() +
    labs(x = "Percent On Track/Mastered", y = "Number of Districts")
```

&nbsp;

How does subgroup performance hang with All Students performance? (It doesn't as much as you'd think.)

``` {r, fig.width = 7, fig.height = 7}
all <- system_numeric %>%
    filter(subgroup == "All Students") %>%
    filter(valid_tests >= 30) %>%
    select(system, subject, OM_percentile) %>%
    rename(OM_percentile_all = OM_percentile)

system_numeric %>%
    filter(!(subgroup %in% c("All Students", "Super Subgroup"))) %>%
    filter(valid_tests >= 30) %>%
    full_join(all, by = c("system", "subject")) %>%
    filter(!is.na(OM_percentile)) %>%
    ggplot(aes(x = OM_percentile_all, y = OM_percentile)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    facet_grid(subject ~ subgroup) +
    theme_hc() +
    labs(x = "All Students On Track/Mastered Percentile", y = "Subgroup On Track/Mastered Percentile")
```

&nbsp;

Movement in percentile ranks in HS Math by subgroup...

```{r, fig.width = 7, fig.height = 4}
system_numeric %>%
    filter(subject == "HS Math") %>%
    filter(!(subgroup %in% c("All Students", "Super Subgroup"))) %>%
    filter(valid_tests >= 30) %>%
    ggplot(aes(x = PA_percentile_prior, y = OM_percentile, colour = Math)) +
    geom_point() +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    facet_grid(. ~ subgroup) +
    theme_hc() +
    labs(x = "2015 Math Proficient/Advanced Percentile", y = "2016 Math On Track/Mastered Percentile")
```

&nbsp;

Movement in percentile ranks in HS English by subgroup...

```{r, fig.width = 7, fig.height = 4}
system_numeric %>%
    filter(subject == "HS English") %>%
    filter(!(subgroup %in% c("All Students", "Super Subgroup"))) %>%
    filter(valid_tests >= 30) %>%
    ggplot(aes(x = PA_percentile_prior, y = OM_percentile)) +
    geom_point() +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    facet_grid(. ~ subgroup) +
    theme_hc() +
    labs(x = "2015 English Proficient/Advanced Percentile", y = "2016 English On Track/Mastered Percentile")
```
